---
title: 'Assignment 2: Data visualization'
author: "Tamas Nagy"
output: html_document
editor_options: 
  chunk_output_type: console
---

You will have to create 3 plots based on the datasets and instructions detailed below. You will find the plots themeselves in the `assignments/assignment_2_plots`. Your task is to write the code that will reproduce the plots as closely as possible.

# Skills needed to solve this assignment

-   Using R and RStudio, reading data
-   Reporting using RMarkdown
-   Using Git and Github (for submitting the task)
-   Data manipulation (e.g. dplyr, tidyr), and working with factors (forcats)
-   Data visuzlization (ggplot2)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(tidytuesdayR)
library(forcats)
```

## Task 1: Climbing expeditions

The 2020-09-22 TidyTueday datasets are about climbing expeditions. From the three datasets, use the "expeditions". Reproduce the plot below! Notice a few things:

-   Use `forcats::fct_lump()` to get the 15 most frequent peaks, and drop the "Other" category.
-   The bars are ordered by the sum of all expeditions (use `fct_reorder()`).
-   The bar colors use the viridis palette and light theme.

```{r}
#Loading data
climber.all <- tt_load_gh("2020-09-22")
climber.expeditions <- tt_download(climber.all, files = climber.all[3])
climber.expeditions <- data.frame(climber.expeditions)
head(climber.expeditions)

# get the 15 most popular

peak.freq <- forcats::fct_lump(climber.expeditions$expeditions.peak_name, n = 15) %>% table()
peak.freq <- data.frame(peak.freq)

# drop 'other'

peak.freq <- peak.freq[-16,]

#Vector for subsetting
names(peak.freq)[names(peak.freq) == "."] <- "Pname"
peak.freq.vec <- as.vector(peak.freq$Pname)

# season freq -> subset 15 most popular
season.freq <- rename(count(climber.expeditions, expeditions.peak_name, expeditions.peak_name, expeditions.season), Freq = n)

plot.base <- season.freq[season.freq$expeditions.peak_name %in% peak.freq.vec,]

# build the plot

#Setting light theme & viridis

theme_set(theme_light())
library(viridis)


 ggplot(plot.base, 
        aes(x = reorder(expeditions.peak_name, Freq, FUN = sum), y = Freq, fill=expeditions.season)) + 
   geom_bar(stat = "identity",width = 0.8) + 
   labs(y="Number of Expeditions", y = "") +
   ggtitle("The 15 most popular peaks stacked by season of expedition") +
   coord_flip() + 
   theme(legend.position = "bottom", axis.title.y = element_blank()) +
   scale_fill_viridis_d(guide_legend(title="Season"))
```

## Task 2: PhDs awarded

The 2019-02-19 TidyTueday dataset is about phd-s awarded by year and field. There is only one dataset, it is called `phd_by_field`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all phd-s by broad fields.
-   To make the x axis breaks pretty, use `scales::pretty_breaks()`, to make the y axis labels comma formatted, use `scales::comma_format()`.
-   The line size is 1.2, the colors are from the brewer "Dark2" palette. The theme is set to minimal.

```{r}
phd <- tt_load_gh("2019-02-19")
phd <- data.frame(tt_download(phd, files = phd))

#aggregate data
phd.agg <- phd[,c(1,4,5)]
phd.agg <- aggregate(.~phd_by_field.broad_field+phd_by_field.year, phd.agg, sum)


#plot
theme_set(theme_minimal())
base.plot <- ggplot(phd.agg,
       aes(x = phd_by_field.year, y = phd_by_field.n_phds, group = phd_by_field.broad_field, color = phd_by_field.broad_field)) +
  geom_line(size = 1.2)

require(scales)

base.plot +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  scale_y_continuous(labels = comma_format(decimal.mark = ",")) +
  scale_color_brewer(palette = "Dark2",guide_legend(title="Broad Field")) +
  ggtitle("Number of Phds in the US by year")
                  
                     

```

## Task 3: Commute in the US

The 2019-11-05 TidyTueday dataset is about commuting to work in each city in the US by bike or on foot. There is only one dataset, it is called `commute`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all commutes by state.
-   Both axis scales are log transformed and the labels comma formatted, using `scales::comma_format()`
-   The point size is 2, . The theme is set to light.

```{r}
commute <- tt_load_gh("2019-11-05")
commute <- data.frame(tt_download(commute, files = commute))

#aggregate data
commute.aggr <- commute[c(8,4,5,9)]
# adding an NA level o factors to keep NAs during aggregate

nm1 <- setdiff(names(commute.aggr[-3]), 'count')
commute.aggr[nm1] <- lapply(commute.aggr[nm1], addNA)
commute.aggr <- aggregate(.~commute.state_abb+commute.mode+commute.state_region, commute.aggr, na.action = na.pass, sum)

#setting up the plot
theme_set(theme_light())
plot.commute <- commute.aggr %>% 
  spread(key = commute.mode, value = commute.n) %>% 
  ggplot(aes(x = Walk,
             y = Bike,
             label = commute.state_abb,
             group = commute.state_region,
             color = commute.state_region)) +
  scale_x_continuous(labels = comma_format(decimal.mark = ","),
                     trans = "log10",
                     limits = c(3000, 500000),
                     breaks = c(3000, 10000, 30000, 100000, 300000)) +
  scale_y_continuous(labels = comma_format(decimal.mark = ","),
                     trans = "log10",
                     limits = c(100, 150000),
                     breaks = c(100, 1000, 10000, 100000)) +
  geom_point(size = 2) + 
  geom_text() + 
  guides(fill=guide_legend(title="New Legend Title")) +
  labs(x="Number of Ppl walking to work (log N)",
       y = "Number of Ppl biking to work (log N)") +
  ggtitle("Number of People walking vs. biking to work in each US state")

plot.commute$labels$colour <- "State Region"

plot.commute
```